---
title: "Text Analysis with R"
author: "Becky Tang and Graham Tierney"
date: "October 3, 2019"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "seahorse"
    fonttheme: "structurebold"
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(tidytext)
library(stringr)
library(tidyr)
library(scales)
library(textdata)
library(wordcloud)
library(reshape2)
library(janeaustenr)
```
# Introduction 
 

## What is text analysis?
- As statisticians and data scientistis, why restrict ourselves to numeric data?
- We can use text as data and employ statistical methods to derive patterns or trends 
  - Example: Given some song lyrics, can we determine who the singer is? What kinds of topics/sentiments does Lizzo like to sing about?
  
## Tidytext
- So far in class, you have been employing tidy data principles, and we will continue to do so here!
- Specifically, we will work in the 'tidy text format': a table with one \textbf{token} per row
```{r introduce_text}
pp <- austen_books() %>%
  filter(book == "Pride & Prejudice") %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  select(-"book")
pp %>% slice(1:11)
```
- How can we make this text into a format we can work with?

## Tokenizing
- A \textbf{token} is a meanginful unit of text. For us, that will be a single word
- Tokenizing is simply splitting a body of text into tokens, which can achieved using the function `unnest_tokens()`

```{r unnest-token, echo = T}
tidy_pp <- pp %>%
  unnest_tokens(word, text)
tidy_pp
```

## Tokenizing cont.
- With the `unnest_tokens()` function, we can easily format any body of text into a user-friendly data frame
- First argument: name of column/variable that text is being unnested into
- Second argument: name of column/variable that is to be unnested 
- Other details:
  - All other columns kept
  - Punctuation removed
  - Defaults is to convert tokens to lowercase

# Twitter Data