---
title: "Text Analysis with R"
author: "Becky Tang and Graham Tierney"
date: "October 3, 2019"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "seahorse"
    fonttheme: "structurebold"
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(tidytext)
library(stringr)
library(tidyr)
library(scales)
library(textdata)
library(wordcloud)
library(reshape2)
library(janeaustenr)
library(lubridate)
library(scales)

```

# Introduction 
 

## What is text analysis?
- As statisticians and data scientistis, why restrict ourselves to numeric data?
- We can use text as data and employ statistical methods to derive patterns or trends 
  - Example: Given some song lyrics, can we determine who the singer is? What kinds of topics/sentiments does Lizzo like to sing about?
  
## Tidytext
- So far in class, you have been employing tidy data principles, and we will continue to do so here!
- Specifically, we will work in the 'tidy text format': a table with one \textbf{token} per row
```{r introduce_text}
pp <- austen_books() %>%
  filter(book == "Pride & Prejudice") %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  select(-"book")
pp %>% slice(1:11)
```
- How can we make this text into a format we can work with?

## Tokenizing
- A \textbf{token} is a meanginful unit of text. For us, that will be a single word
- Tokenizing is simply splitting a body of text into tokens, which can achieved using the function `unnest_tokens()`

```{r unnest-token, echo = T}
tidy_pp <- pp %>%
  unnest_tokens(word, text)
tidy_pp
```

## Tokenizing cont.
- With the `unnest_tokens()` function, we can easily format any body of text into a user-friendly data frame
- First argument: name of column/variable that text is being unnested into
- Second argument: name of column/variable that is to be unnested 
- Other details:
  - All other columns kept
  - Punctuation removed
  - Defaults is to convert tokens to lowercase

# Twitter Data

## Democratic candidate tweets
- Time to work with some fun data!
- We have pulled tweets from four Democratic candidates: Joe Biden, Kamala Harris, Bernie Sanders, and Elizabeth Warren
```{r}
setwd("~/Documents/data_expedition2019/data")
tweets <- read.csv("dem_cand_tweets.csv")
#colnames(tweets)
tweets <- tweets %>%
  mutate(timestamp = ymd_hms(created_at)) %>%
  select("created_at", "screen_name", "text", "is_retweet", "retweet_count", "timestamp")
tweets %>%
  select("screen_name", "text") %>%
  slice(1:2)
```

## Tidy tweets
```{r}
remove_reg <- "&amp;|&lt;|&gt;"
tidy_tweets <- tweets %>%
  #get rid of retweets
  filter(!str_detect(text, "^RT")) %>%
  mutate(text = str_remove_all(text, remove_reg))%>%
  unnest_tokens(word, text, token = "tweets")
head(tidy_tweets)
```

## Word counts
- What are the most commonly tweeted words by these candidates?
```{r}
tidy_tweets %>%
  count(word, sort = T)
```

## Stop words
```{r, echo = T}
data("stop_words")
head(stop_words)
tidy_tweets <- tidy_tweets %>%
  filter(!word %in% stop_words $ word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]"))
tidy_tweets %>%
  count(word, sort = T)
tidy_tweets %>%
  group_by(screen_name) %>%
  count(word, sort = T)
```

## Word frequency
- While it's nice to get raw counts, but we may prefer to know which words are used more often \textit{relative} to other words. For this we can look towards word frequencies:

```{r frequency}
tidy_tweets %>%
  count(word, sort = T) %>%
  mutate(freq = n / n()) %>%
  head()

frequency <- tidy_tweets %>%
  group_by(screen_name) %>%
  count(word, sort = T) %>%
  left_join(tidy_tweets %>%
              group_by(screen_name) %>%
              summarise(total = n()),
            by = "screen_name") %>%
  mutate(freq = n / total)

frequency %>%
  head()
```


## Plot of word frequency
- We can visualize the differences and similarities in word frequencies between two candidates: 
```{r frequency-plot, warning=FALSE}
# make into nice plot
frequency <- frequency %>%
  select(screen_name, word, freq) %>%
  spread(screen_name, freq) %>%
  arrange(BernieSanders, ewarren, JoeBiden, KamalaHarris)

# choose 2 candidates
ggplot(frequency, aes(JoeBiden, ewarren))+
  # hide discreteness
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = T, vjust = 1.5)+
  scale_x_log10(labels = percent_format())+
  scale_y_log10(labels = percent_format())+
  geom_abline(color = "blue")
```
- Words near the blue line are used with roughly equal frequencies between Joe Biden and Elizabeth Warren
- Words far away from the blue line are favored much more by one candidate than the other

## Log-odds ratio
- Looking beyond raw frequencies, we can now begin to investigate which candidate is more likely to use a given word in a tweet
- To do so, we will utilize the \textbf{log-odds ratio}
$$\text{OR}_{A:B} = \dfrac{\text{odds}_A}{\text{odds}_B} = \dfrac{\text{prob}(word)_A}{\text{prob}(word)_B}$$
\begin{align*}
\log \text{OR}_{A:B} = \log \left( \dfrac{\text{prob}(word)_A}{\text{prob}(word)_B}\right) &= \log \left( \dfrac{n_A / total_A}{n_B / total_B}\right) \\ 
& \approx \log \left( \dfrac{(n+1)_A / (total+1)_A}{(n+1)_B / (total+1)_B}\right)
\end{align*}

## Word usage

```{r word usage}
candidates <- c("JoeBiden", "KamalaHarris")
word_ratios <- tidy_tweets %>%
  filter(screen_name %in% candidates) %>%
  filter(!str_detect(word, "^@")) %>%
  count(word, screen_name) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(screen_name, n, fill = 0) %>%
  mutate_if(is.numeric, list(~(. +1) / sum(.) + 1)) %>%
  mutate(logratio = log(eval(parse(text = candidates[1])) / eval(parse(text = candidates[2])))) %>%
  arrange(desc(logratio))

```
